Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-10-15 exp1/present2013.tex
cut fractionals
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@679 43ef1190-90fd-4dd7-a769-d0121f02b80e

  \begin{block}{Message for the takeaway}
    If you develop techniques for performance enhancements of complex
    systems, you need to think carefully about how you prove that 
    you've achieved something of general validity. 
  \end{block}

@@

\begin{itemize}
\item Large number of parameters to test for complex scenarios.
\item No structured approach to investigate flaws.
\item No meaningful summary of the overall performance.
\item Standardized benchmarks cannot test assertions outside of the
  standard.
\item Attempting to neutralize the effect of certain optimization
  techniques oversimplifies the test.
\end{itemize}
\end{frame}
\note{  
  \begin{itemize}
  \item If you standardize hardware, you can't test claims about
  improved hardware. You'd need interactions too
  \item Caching is a good practice, and its effect need to be a part
    of the test, not gotten rid of.
  \item That said benchmarking is still nice to have if you know what
    queries you will run, it can't replace that.
\end{itemize}

@@

\note{
  \begin{itemize}
  \item Unreplicated, so we don't have a standard error, so Lenth used
    a restricted subset the median of the effects to create a Pseudo
    Standard Error
  \item Don't worry about the axes
  \item Yes, the Implementation is better, we knew that, 
  \item Interactions with Implement with joins are
    naturally strong
  \item ``Union'', ``Optional'' and ``Lang'' demanding things
  \item ``BGPComp:Lang'' and ``Lang:Union'' interactions might be due
    to that many more triples need to be searched for a language tag
    in one of the levels of the interacting factors.
  \item But what about ``Implement::TripleC''? Leave that hanging for
    a bit.
  \item We want to do a hypothesis test
  \item Two factors do not contribute significantly to the variation
    ``Range'' and ``Machine''
  \end{itemize}
}

@@

\end{frame}
\note{The paper shows how similar testing can be done to investigate
  when the new implementation fails.}

\subsection{Fractional Factorial Experiments}

\begin{frame}{Experimental Economy}

  \begin{itemize}
  \item Full factorial experimental goes as $2^n$.
  \item SPARQL Endpoint evaluation is inherently complex, with many
    possible factors.
  \item May reduce the size of the experiment with Fractional
    Factorial Experiments.
  \item The cost is explanatory power due to \emph{aliasing}.
  \item E.g. we cannot tell the difference of an effect of
    ``TripleC::BGPComp'' from ``Machine::Range''
  \item Easy in R to declare which effects must not be aliased.
  \item We have made two designs: One 32-run and one 64-run.
  \end{itemize}

\end{frame}
\note{ That is to say, a detected increase in
run time from larger Basic Graph Patterns for large databases, can
also be explained by a less powerful machine that is worse at
evaluating FILTER clauses with ranges. }

\begin{frame}{32-run experiment}
\begin{figure}[ht!]
  \centerline{%
    \includegraphics[height=0.85\textheight,width=.9\textwidth]{frac32normal.pdf}}
  \label{fig:frac32normal}
\end{figure}

\end{frame}
\note{

  \begin{itemize}
  \item The total variance is great, due to few runs, so few
    conclusions can be drawn
  \item Implementation is good, and Union is hard
  \item In the paper, we also do a hypothesis test that returns $p =
    0.00098$. Still very low
  \item Very rough conclusions; useful in a CI system
  \end{itemize}
}

\begin{frame}{64-run experiment}
\begin{figure}[ht!]
  \centerline{%
    \includegraphics[height=0.85\textheight,width=0.96\textwidth]{frac64normal.pdf}}
\end{figure}

\end{frame}
\note{Have two Lenth criterions, $\alpha=0.15$ and crosses at
  $\alpha=0.05$. ($\alpha$ is the probability of rejecting $H_0$ when
  it is true)}

\begin{frame}{Analysis of 64-run experiment}

  \begin{itemize}
  \item Find most of the same significant effects as the full experiment.
  \item However, we see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$.
  \item ``Machine'', ``TripleC'' and ``Range'' are inactive, we can
    perform hypothesis test as above, and find $p = 5.9 \cdot 10^{-6}$.
  \end{itemize}
\end{frame}
\note{
  \begin{itemize}
  \item We see that ``Implement:TripleC'' emerges only for
    $\alpha=0.15$, which is important to improve when designing a
    study for actual tests.
  \item We have shown that how the experiment could be set up, we have
    shown the analysis and a hypothesis test.
  \item Should this simplistic experiment stand up to scrutiny?
  \end{itemize}
}

>>>>

  \begin{block}{Message for the takeaway}
    If you develop techniques for performance enhancements of complex
    systems, you need to think carefully about how you prove that 
    you've achieved something of general validity. This is my proposal
    in that direction.
  \end{block}

@@

\begin{itemize}
\item Large number of parameters to test for complex scenarios.
\item No structured approach to investigate flaws.
\item No meaningful summary of the overall performance.
\item Standardized benchmarks cannot test assertions outside of the
  standard.
\item Attempting to neutralize the effect of certain optimization
  techniques oversimplifies the test.
\item \emph{A benchmark is an engineering tool, not a scientific experiment.}
\end{itemize}
\end{frame}
\note{  
  \begin{itemize}
  \item If you standardize hardware, you can't test claims about
  improved hardware. You'd need interactions too
  \item Caching is a good practice, and its effect need to be a part
    of the test, not gotten rid of.
  \item That said benchmarking is still nice to have if you know what
    queries you will run, it can't replace that.
\end{itemize}

@@

\note{
  \begin{itemize}
  \item Those that are deemed significant are labelled.
  \item Don't worry about the axes
  \item Yes, the Implementation is better, we knew that, 
  \item Interactions with Implement with joins are
    naturally strong
  \item ``Union'', ``Optional'' and ``Lang'' demanding things
  \item ``BGPComp:Lang'' and ``Lang:Union'' interactions might be due
    to that many more triples need to be searched for a language tag
    in one of the levels of the interacting factors.
  \item But what about ``Implement::TripleC''? Leave that hanging for
    a bit.
  \item We want to do a hypothesis test
  \item Two factors do not contribute significantly to the variation
    ``Range'' and ``Machine''
  \end{itemize}
}

@@

\end{frame}
\note{  
  \begin{itemize}
  \item The paper shows how similar testing can be done to investigate
  when the new implementation fails.} 
\item The paper has two fractional factorial experiments.
  \item Should this simplistic experiment stand up to scrutiny?
  \end{itemize}
}
